Sell Prevention and Buy Cure: An Exploration of Indirect Conflicts

### 2014-01-29 iced-2013/paper.tex
ICED peggy and casper changes.

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, and scalability. To better understand these issues, 
we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
source projects, and a survey with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
we obtained insights about events and conditions in which indirect conflicts are likely to occur in
software development, that developers prefer to use detection and resolution processes or tools
over those of prevention, that they do not want awareness mechanisms which provide non actionable results,
and that there exists a gap in software evolution analytical tools.

@@

Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts involve source code change influences such as making a source code change 
which influences behavior negatively in another location in the system. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

@@

We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts.

@@

We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
addressed our 3 main research questions. Second, a survey was conducted
which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
interview participants to once again confirm our insights.
We used grounded theory techniques to analyze the information provided from all three data gathering stages.

@@

Our interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}: IBM, Mozilla, The GNOME Project, Microsoft Corporation, Subnet Solutions, Ruboss Technology Corporation, 
Amazon, Exporq Oy, Kano Apps, Fireworks Design, James Evans and Associates, and Frost Tree Games. 
Our participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
with an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management
at some capacity. See Table~\ref{tab:demo} for more demographic details.

@@

Participants were invited to be interviewed by email and were sent a single reminder email one week
after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use.

@@

Interview participants first answered a number of  demographic questions. We then asked them to describe various software development experiences regarding our three research questions.
Specifically, ten semi-structured  topics directly related to our research questions guided our interview:

%Our three questions were studied by having the interview participants talk about their experiences and opinions in
%\totalCategories{} .

\begin{itemize}

\item Software development tools for dependency tracking and awareness.

\item Software development process for preventing indirect conflicts.
\item Software artifact dependency levels and where conflicts can arise.

\item How developers find internal or external software dependencies.

\item Examples of indirect conflicts from real world experiences.
\item How indirect conflicts are detected and found.

\item How indirect conflict issues are solved or dealt with.
\item Developer opinion of preemptive measures to prevent indirect conflicts.

\item Developer opinion on what types of changes are worth a preemptive action.

\item Developer opinion on who is responsible for fixing or preventing indirect conflicts.
\end{itemize}

While each of the \totalCategories{} topics had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time.

%We account for these non answers
%in our analysis and results as seen in Section~\ref{sec:results}.

\subsection{Survey Participants}

Our survey respondents were different from our interviewees. We invited our survey participants from a similar breadth of open and closed source software development 
companies and projects as the interviews participants with two main exceptions. The software organizations
that remained the same between interview and survey were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon.
Participants who took part in the round of interviews were not invited to the survey but were asked to act as a  point of contact for other 
developers in their team, project, or organization who may be interested
in completing the survey. Further, two other groups of developers were asked to
participate as well, these being GitHub users as well as Apache Software Foundation (Apache) developers. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

\subsection{Survey Procedure}

Survey participants were invited to participate in the survey by email. No reminder email was sent as the
survey responses were not connected with the invitation email addresses and thus participants who did respond
could not be identified. We directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. We attribute the low response rate with: the surveys
were conducted during the months of July and August while many participants may be away from their regular positions.
and our GitHub and Apache participants could not be verified as to whether or not they actively support the
email addresses used in the invitations. In addition, the survey was considered by some to be long and require
more development experience than may have been typical of some of those invited to participate.

The  survey we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar
or different opinions from those already acquired from the interviews. The survey went through two rounds of
piloting. Each pilot round consisted of five participants, who were previously interviewed, completing the survey
with feedback given at the end. Not only did this allow us to create a more polished survey, but it also allowed 
the previously interviewed developers to examine the insights we developed.
%Each non demographic question was made optional as it was shown 
%through the interviews that some questions require more experience from participants than may be provided.

\subsection{Data Analysis}
To analyze both the interview and survey data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data. For an exploratory study such as
the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and surveys (when open ended questions were involved) was analyzed using open and axial coding. Open coding involves
assigning codes to what participants said at a low sentence level or abstractly at a paragraph or full answer level. These
codes were defined as the study progressed and different hypotheses began to grow. We finally use axial coding it order to
link our defined codes through categories of grounded theory such as context and consequences. 
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
3 criteria that are commonly used in evaluating grounded theory studies.

@@

We now present our results of both the interviews and surveys conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
results from which we draw our discussion to be seen in Section~\ref{sec:disc}.

\subsection{What events or conditions lead to indirect conflicts?}

From the interviewed participants, 12 developers believe that a large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
In light of object contracts, 14 interviewed developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contacts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

In regards to object contract changes, 9 developers currently working with large scale database applications listed database
schemas as a large source of indirect conflicts while 5 developers that work on either software
library projects or are in test said that methods or functions were the root of their indirect conflict issues. 
7 interviewed developers mentioned that indirect conflicts occur when a major update to an external project, 
library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
other notable indirect conflict artifacts were user interfaces in web development and full components in component 
base game architecture.

From the interviewed participants, 11 explained that indirect conflicts occur ``all the time'' in
their development life cycle with a minimum occurrence of once a week, with more serious issues tending
to occur once a month. 65\% of surveyed developers answered that indirect conflicts occur on minimum bi-weekly,
with the majority of developers saying that weekly occurrences are most common.

12 developers interviewed said that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project 
changes a lot, especially with agile''. Surveyed developers also added ``indirect conflicts after a release 
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a 
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during 
large revamps or the implementation of cross-cutting new features.''. Surveyed participants also answered that 
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c | c | c |}
\hline
Occurrences & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
\hline
\hline
Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
After the first release & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
Late stages of development & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of survey questions to how often indirect conflicts occur, in terms of percentage
of developers surveyed.\label{tab:often}}
\end{table*}

In terms of organizational structure, surveyed participants answered that as a project becomes larger and more 
developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
saying they are like to occur in single developer projects.

\subsection{What mitigation techniques are used by developers in regards to indirect conflicts?}

In terms of preventative processes used for indirect conflicts, 3 major components were found.
First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or understand
when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
contract when possible, and that an object's contract is used as a type of documentation towards awareness of the 
software object. One developer stated that ``design by contract was invented to solve this problem and it does it 
quite well'', while another noted that software object contracts do solve the problem in theory, but that doesn't 
mean that problems don't still occur in practice.
Second, 21\% of interview developers mentioned that the add and deprecate 
model is used to prevent indirect conflicts once
the project, feature, or component has reached a stable or mature point.
Add and deprecate meaning instead of editing code, the developer simply clones old code (if needed), and edits the clone
while slowly phasing out the out the old code in subsequent releases or ad needed. This allows other software to 
use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
Lastly, pure developer experience was mentioned with 7 developers mentioning that when planning code changes, 
either a very experienced member of the project was involved in the planning and has duties to foresee any 
indirect conflicts that may arise, or that developers must use their personal knowledge to predict where indirect 
conflicts will occur while implementing.

Of the the 37 developers surveyed who could give positive identification of preventative processes for indirect conflicts,
27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
by contract or the testing of a methods contract, and 14\% said that add and
deprecate was used in their projects to avoid indirect conflicts.

In regards to catching indirect conflicts, 13 interviewed developers mentioned forms of testing (unit, and integration) 
as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration 
tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
not all indirect conflicts should be caught. 31\% of surveyed developers said build processes (either nightly builds or building the project 
themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run 
time testing. The surveyed developers confirmed these results with 49\% mentioning forms of testing as the major tool used to 
catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
Surveyed developers also mentioned review process and personal expertise as factors of catching indirect conflicts.

Once an indirect conflict has occurred and developers need to resolve it, 14 developers interviewed said
they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time 
frames of when events occurred. Of the developers surveyed, 23\% said they used native IDE tools and 21\% said they use 
features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers 
mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading 
of formal documentation.

@@

\subsection{What do developers want from future indirect conflict tools?}

@@

In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
are already in place to catch potential issues such as testing, code review, 
individual knowledge, or static language analysis tools. 68\% of surveyed developers said that they would always
want to be notified about method signature changes as they have a high chance to break the code as opposed to only
23\% who always wanted to be notified on a pre or post condition change and 27\% who want to be notified for a user
interface change. Other change types varied from never being notified to most times being notified, showcasing the
complexity of change types which may or may not negatively affect a project.

When asked about curative tools, developers could only suggest that resolution times be decreased by different means.
Surveyed developers suggested the following improvements to curative tools:

@@

@@

Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue'', meaning
that the developer only wants the awareness if the item will require action on their part to resolve the issue.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload. 
Developers simply do not want to know about events which effect them, but require no action on their part.

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption. 
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for 
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible. \textit{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

>>>>

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, and scalability. To better understand these issues, 
we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
source projects, and a questionnaire with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis, 
we obtained insights about events and conditions in which indirect conflicts are likely to occur in
software development, that developers prefer to use detection and resolution processes or tools
over those of prevention, that they do not want awareness mechanisms which provide non actionable results,
and that there exists a gap in software evolution analytical tools.

@@

Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts (ICs) are caused by source code changes that negatively impact another 
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influence negatively. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

@@

We interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory questionnaire of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. Our findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and there exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
context of indirect conflicts.

For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose 
recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
Kim conducted several focus groups as well as web questionnaires to determine these results. These areas of interest resonate with what was
found in this paper. Developers want to know when a code change is going to interfere with their work in a potentially negative way and
require action on the developer's half.  Kim
also found that developers are concerned with interfaces of objects and when those interfaces change, similar to the object contracts that
were found in this paper. Finally, Kim also identified the same issues towards information overload through false positives with developers
noting ``I get a big laundry list... I see the email and I delete it''.

@@

We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which 
addressed our 3 main research questions. Second, a questionnaire was conducted
which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
interview participants to once again confirm our insights.
We used grounded theory techniques to analyze the information provided from all three data gathering stages.

@@

Our interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
Our participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience had an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management
at some capacity.

@@

Participants were invited to be interviewed by email and were sent a single reminder email one week
after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible, or over video chat when not possible, and recorded for audio content only.

@@

Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences regarding our three research questions.
Specifically, ten semi-structured questions related to our research questions where ask and guided our interview:

\begin{itemize}
\item What tools are used for dependency tracking?
\item What processes are used for preventing ICs?
\item What artifact dependency levels are analyzed and where can conflicts arise?
\item How are software dependencies found?
\item Give examples of ICs from real world experiences.
\item How are ICs detected or found?
\item How are IC issues solved or dealt with?
\item Opinion of preemptive measures to prevent ICs.
\item What types of changes are worth a preemptive action?
\item Who is responsible for fixing or preventing ICs?
\end{itemize}

Not all participants had strong opinions or any experience on every category listed above. For these participants, answers 
to the specific categories were not required. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time.

\subsection{Questionnaire Participants}

Our questionnaire respondents were different from our interviewees. We invited our questionnaire participants from a similar 
breadth of open and closed source software development 
companies and projects as the interviews participants with two main exceptions, the addition of GitHub users as well as Apache Software Foundation (Apache) developers.
The software organizations that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation, 
Subnet Solutions, and Amazon. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

\subsection{Questionnaire Procedure}

Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
were conducted during the months of July and August while many participants may be away from their regular positions.

The questionnaire we designed ~\footnote{http://thesegalgroup.org/people/jordan-ell/iced\_survey/}
was based on the insights we obtained from the round of interviews, and was intended to confirm some of these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews. The question topics asked in the questionnaire were:

\begin{itemize}
\item What frequency do ICs occur at around different project milestones?
\item How does team size affect IC frequency?
\item What software change types do developers care about?
\item What processes are used for preventing ICs?
\item What tools are used for detecting ICs?
\item What tools are used for debugging ICs?
\end{itemize}

\subsection{Data Analysis}
To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data (although this paper only uses grounded
theory techniques for analysis and not for a final theory). For an exploratory study such as
the ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding.
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using 
3 criteria that are commonly used in evaluating grounded theory studies.

@@

We now present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. We restate each research question, followed by our quantitative and qualitative
results from which we draw our discussion to be seen in Section~\ref{sec:disc}. In each subsection, quantitative data
given refers to interviews conducted unless explicitly said otherwise.

\phantomsection
\addcontentsline{toc}{subsection}{RQ1}
\subsection*{\normalfont{RQ1.} \it{What events or conditions lead to indirect conflicts?}}

12 developers believe that a large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract. Object contracts are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed
to work; made famous by Eiffel Software's~\footnote{http://www.eiffel.com/} ``Design by Contract''\texttrademark. 
In light of object contracts, 14 developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers 
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contacts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

In regards to object contract changes, 9 developers currently working with large scale database applications listed database
schemas as a large source of indirect conflicts while 5 developers that work on either software
library projects or are in test said that methods or functions were the root of their indirect conflict issues. 
7 developers mentioned that indirect conflicts occur when a major update to an external project, 
library, or service occurs with one developer noting ``their build never breaks, but it breaks ours''. Some
other notable indirect conflict artifacts were user interfaces in web development and full components in component 
base game architecture.

11 developers explained that indirect conflicts occur ``all the time'' in
their development life cycle with a minimum occurrence of once a week, with more serious issues tending
to occur once a month. To confirm this, 64\% of questionnaire participants answered that indirect conflicts occur on minimum bi-weekly,
with the 25\% saying that weekly occurrences are most common (seen in ~\ref{tab:often}). 5 questionnaire participants
also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
differ.

12 developers said that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. Developers said ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project 
changes a lot, especially with agile''. Questionnaire participants also added ``indirect conflicts after a release 
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a 
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during 
large revamps or the implementation of cross-cutting new features.'' in order to confirm our interview results. 
Questionnaire participants also answered that 
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c | c | c |}
\hline
Occurrences & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
\hline
\hline
In general & 18\% & 25\% & 21\% & 16\% & 3\% & 5\% & 11\% \\ \hline
Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
After the first release & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
Late stages of development & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of questionnaire as to how often indirect conflicts occur, in terms of percentage
of questionnaire participants.\label{tab:often}}
\end{table*}

In terms of organizational structure, questionnaire participants answered that as a project becomes larger and more 
developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
occur. However, indirect conflicts still occur at a lower number of developers as well with even 43\% of developers
saying they are like to occur in single developer projects. This can be seen in Table~\ref{tab:env}.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c |}
\hline
Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
\hline
\hline
Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
Developing in a team between 2 - 5 developers (inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
Developing in a multi team environment (inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
\end{tabular}
\end{center}
\caption{Questionnaire results about development environments in which indirect conflicts are likely to occur, in terms of percentage
of questionnaire participants.\label{tab:env}}
\end{table*}

\phantomsection
\addcontentsline{toc}{subsection}{RQ2}
\subsection*{\normalfont{RQ2.} \it{What mitigation techniques are used by developers in regards to indirect conflicts?}}

In terms of preventative processes used for indirect conflicts, 3 major components were found.
First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or understand
when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
contract when possible, and that an object's contract is used as a type of documentation towards awareness of the 
software object. One developer stated that ``design by contract was invented to solve this problem and it does it 
quite well'', while another noted that software object contracts do solve the problem in theory, but that doesn't 
mean that problems don't still occur in practice.
Second, 21\% of developers mentioned that the add and deprecate 
model is used to prevent indirect conflicts once
the project, feature, or component has reached a stable or mature point.
Add and deprecate meaning instead of editing code, the developer simply clones old code (if needed), and edits the clone
while slowly phasing out the out the old code in subsequent releases or ad needed. This allows other software to 
use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
Lastly, pure developer experience was mentioned with 7 developers mentioning that when planning code changes, 
either a very experienced member of the project was involved in the planning and has duties to foresee any 
indirect conflicts that may arise, or that developers must use their personal knowledge to predict where indirect 
conflicts will occur while implementing.

Of the the 37 questionnaire participants who could give positive identification of preventative processes for indirect conflicts,
27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
by contract or the testing of a methods contract, and 14\% said that add and
deprecate was used in their projects to avoid indirect conflicts. This is a confirmation of what was found in the interviews.

In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration) 
as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration 
tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
not all indirect conflicts should be caught. In corroboration, 31\% of questionnaire participants said build processes (either nightly builds or building the project 
themselves), and others mentioned code reviews while those dealing with a user interface mentioned user complaints from run 
time testing. The questionnaire participants confirmed these results with 49\% mentioning forms of testing as the major tool used to 
catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.

Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time 
frames of when events occurred. Of the questionnaire participants, 23\% said they used native IDE tools and 21\% said they use 
features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers 
mentioned a form of communication with other developers in aid to solving these conflicts and only 4\% mentioned the reading 
of formal documentation.

@@

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c |}
\hline
Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
\hline
\hline
Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
\end{tabular}
\end{center}
\caption{Questionnaire results about source code changes that developers deem notification worthy, in terms of percentage
of questionnaire participants.\label{tab:pre}}
\end{table*}

\phantomsection
\addcontentsline{toc}{subsection}{RQ3}
\subsection*{\normalfont{RQ3.} \it{What do developers want from future indirect conflict tools?}}

@@

In terms of catching indirect conflicts, interviewed developers suggested that proper software development processes 
are already in place to catch potential issues such as testing, code review, 
individual knowledge, or static language analysis tools. Further, developers said that having strict change type monitoring 
for indirect conflicts does not work due to the inherent complexities of indirect conflicts.
This can be seen confirmed in that questionnaire participants had very few cases in which they always wanted to be
alerted to a change type as seen in Table~\ref{tab:pre}. Only method signature changes caused 68\% of questionnaire participants
to want to be always notified as they have a high chance to break the code. This is opposed to other change types listed in
Table~\ref{tab:pre} which have 27\% or lower of developers always wanting to be notified. This again showcases the
complexity of indirect conflicts through change types which may or may not negatively affect a project.

When asked about curative tools, developers could only suggest that resolution times be decreased by different means.
Interview and questionnaire participants suggested the following improvements to curative tools:

@@

As per our results, three main discussion points have arisen. What developers really want, and don't want from indirect
conflict awareness tools, if indirect conflicts are better handled by extensively trying to prevent them or by simply
fixing them as they occur with less prevention, and a gap in analytical tools for software dependencies found in current
industry settings.

@@

Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
developers only want awareness of an event if the event forces the developer to take some action. In a sense, developers
don't care about what they don't need to account for. ``We would want a high probability of the [reported] change being an issue''.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to 
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload. 
Developers simply do not want to know about events which effect them, but require no action on their part.
